### 2016-02-21 Paper0/ms.tex
more consistent with drizzled

\documentclass[iop, numberedappendix]{emulateapj}
\usepackage{apjfonts}
%\documentclass[12pt,preprint]{aastex}

%\topmargin 0.6in

%\usepackage{deluxetable}
\usepackage{color}
\definecolor{citeRGB}{rgb}{0,0.1,0.7}
\usepackage[hyperfootnotes=true,naturalnames=true,letterpaper,pdfstartview=FitH,pdfpagemode=UseNone,colorlinks=true,citecolor=citeRGB]{hyperref}

%%%% The following code forces Hyperref to only link the year in a citation
%%%% Taken from http://tex.stackexchange.com/questions/23227/how-to-hyperlink-only-the-year-part-when-using-natbib-and-hyperref
% \usepackage{etoolbox}
% 
% \makeatletter
% 
% % Patch case where name and year have no delimiter
% \patchcmd{\NAT@citex}
%   {\@citea\NAT@hyper@{\NAT@nmfmt{\NAT@nm}\NAT@date}}
%   {\@citea\NAT@nmfmt{\NAT@nm}\NAT@hyper@{\NAT@date}}
%   {}% Do nothing if patch works
%   {}% Do nothing if patch fails
% 
% % Patch case where name and year have basic delimiter
% \patchcmd{\NAT@citex}
%   {\@citea\NAT@hyper@{%
%      \NAT@nmfmt{\NAT@nm}%
%      \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
%      \NAT@date}}
%   {\@citea\NAT@nmfmt{\NAT@nm}%
%    \NAT@aysep\NAT@spacechar%
%    \NAT@hyper@{\NAT@date}}
%   {}% Do nothing if patch works
%   {}% Do nothing if patch fails
% 
% % Patch case where name and year are separated by a prenote
% \patchcmd{\NAT@citex}
%   {\@citea\NAT@hyper@{%
%      \NAT@nmfmt{\NAT@nm}%
%      \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
%        {\@citeb\@extra@b@citeb}%
%      \NAT@date}}
%   {\@citea\NAT@nmfmt{\NAT@nm}%
%    \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi%
%    \NAT@hyper@{\NAT@date}}
%   {}% Do nothing if patch works
%   {}% Do nothing if patch fails
% 
% \makeatother

%\gdef\zphot{$z_\mathrm{phot}$}
\gdef\HST{\textit{HST}}
\gdef\G141{\textit{G141}}
\gdef\F140W{\textit{F140W}}
\gdef\fluxcgs{\mathrm{erg\ s^{-1}\ cm^{-2}}}
\gdef\micront{$\mu$m}
\gdef\micronm{\mu\mathrm{m}}
\gdef\aXe{\texttt{aXe}}
\gdef\flux_radius{\textsc{flux\_radius}}
\gdef\epers{\textit{e}$^{-}$ s$^{-1}$}

\gdef\logOH{12 + \log \left(\mathrm{O/H}\right)}
\gdef\UmV{U-V}
\gdef\VmJ{V-J}
\gdef\SFRuvir{SFR_\mathrm{UV+IR}}
\gdef\peryr{\mathrm{yr}^{-1}}
\gdef\kms{km\,s$^{-1}$}
\gdef\mum{$\mu\mathrm{m}$}
\gdef\24mum{$24\,\mu\mathrm{m}$}
\gdef\arcsec{^{\prime\prime}}
\gdef\Jy{\mathrm{Jy}}
\gdef\uJy{\mu\mathrm{Jy}}
\gdef\LIR{L_\mathrm{IR}}
\gdef\LUV{L_\mathrm{UV}}
\gdef\4ang{4000\,\AA}
\gdef\msun{$M_{\odot}$}
\gdef\mathmsun{\ M_{\odot}}
\gdef\mathgyr{\ \textrm{Gyr}}
\gdef\mathmyr{\ \textrm{Myr}}
\gdef\pegase{\textsc{P\'{e}gase}}
\gdef\eazy{\texttt{Eazy}}
\gdef\galfit{\texttt{galfit}}
\gdef\emcee{\texttt{emcee}}
\gdef\interrest{\texttt{InterRest}}
\gdef\fast{\texttt{fast}}
\gdef\zCOSMOS{$z$\textsc{COSMOS}}
\gdef\zspec{z_\mathrm{spec}}
\gdef\zphot{z_\mathrm{phot}}
\gdef\zml{z_\mathrm{ML}}
\gdef\zprior{z_\mathrm{prior}}
\newcommand{\weblink}[1]{\href{#1}{#1}}
\gdef\UDFj{UDFj-39546284}
\gdef\compareID{UDF-40106456}
\gdef\Lya{\mathrm{Ly}\alpha}
\gdef\Halpha{\mathrm{H}\alpha}
\gdef\Hbeta{\mathrm{H}\beta}
\gdef\othree{[\ion{O}{3}]$\lambda$4959+5007}
\gdef\epers{electrons~s$^{-1}$}
\gdef\xxx{\textbf{xxx}}

\gdef\lineflux{$3.5\pm1.3\times10^{-18}\,\fluxcgs$}

\shortauthors{Brammer et al.}
\shorttitle{$K_s$-band imaging of the Frontier Fields}
%\slugcomment{Draft (\today)}
\slugcomment{Submitted to ApJSS}

%\citestyle{aa}

\begin{document}

\title{Ultra-deep $K_s$-band Imaging of the \textit{Hubble} Frontier Fields}

% \footnotetext[*]{Based on observations made with the NASA/ESA \textit{Hubble
% Space Telescope}, programs GO-11640, 12177 and 12328, obtained at the
% Space Telescope Science Institute, which is operated by the Association of
% Universities for Research in Astronomy, Inc., under NASA contract NAS
% 5-26555.}

\author{Gabriel B. Brammer\altaffilmark{1},
Danilo Marchesini\altaffilmark{2},
et~al.}
% Pieter G.\ van Dokkum\altaffilmark{2},
% Garth D.\ Illingworth\altaffilmark{4},
% Rychard J.\ Bouwens\altaffilmark{3},
% Ivo Labb\'e\altaffilmark{3}, 
% Marijn Franx\altaffilmark{3},
% Ivelina Momcheva\altaffilmark{2},
% Pascal A.\ Oesch\altaffilmark{4}}
% Rub\'en S\'anchez-Janssen\altaffilmark{1}, 
% Ivo Labb\'e\altaffilmark{2},
% Elisabete da Cunha\altaffilmark{3},
% Dawn K.\ Erb\altaffilmark{4},
% Marijn Franx\altaffilmark{2},
% Mattia Fumagalli\altaffilmark{2},
% Britt Lundgren\altaffilmark{5},
% Ivelina Momcheva\altaffilmark{5},
% Erica Nelson\altaffilmark{5},
% Shannon Patel\altaffilmark{2},
% Ryan Quadri\altaffilmark{7},
% Hans-Walter Rix\altaffilmark{3},
% Rosalind E.\ Skelton\altaffilmark{5},
% Kasper B.\ Schmidt\altaffilmark{3},
% Arjen van der Wel\altaffilmark{3},
% Pieter G.\ van Dokkum\altaffilmark{5},
% David A.\ Wake\altaffilmark{5},
% Katherine E.\ Whitaker\altaffilmark{5}}

% Rachel Bezanson\altaffilmark{5},
% Xiaohui Fan\altaffilmark{6},
% Natascha F\"orster Schreiber\altaffilmark{8},
% Garth D.\ Illingworth\altaffilmark{9},
% Joel Leja\altaffilmark{5},
% Dan Magee\altaffilmark{9},
% Patrick McCarthy\altaffilmark{11},
% Adam Muzzin\altaffilmark{2},
% Ryan Quadri\altaffilmark{11},
% Charles C.\ Steidel\altaffilmark{12},
% Tomer Tal\altaffilmark{5},
% Anna Williams\altaffilmark{13}

\email{brammer@stsci.edu}

\altaffiltext{1}{ESA/AURA, Space Telescope Science Institute, 3700 San Martin Dr., Baltimore, MD 21218, USA}
\altaffiltext{2}{Physics and Astronomy Department, Tufts University, Medford, MA, 02155, USA}
% \altaffiltext{6}
% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
% \altaffiltext{3}
% {Leiden Observatory, Leiden University, NL-2300 RA Leiden, Netherlands}
% \altaffiltext{4}
% {UCO/Lick Observatory, University of California, Santa Cruz, CA 95064}
% \altaffiltext{3}
% {Max Planck Institute for Astronomy (MPIA), K\"onigstuhl 17,
% 69117, Heidelberg, Germany}
% \altaffiltext{4}
% {Department of Physics, University of Wisconsin-Milwaukee, P.O. Box 413,
% Milwaukee, WI 53201, USA}
% \altaffiltext{5}
% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
% \altaffiltext{7}
% {Carnegie Observatories, 813 Santa Barbara Street, Pasadena, CA 91101, USA}

\begin{abstract}

We present ultra-deep $K_s$-band imaging of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, MACS-0717 and MACS-1149 (\xxx\ include MOSFIRE???), which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 4.5$\,\mum.  \textit{VLT}/HAWKI integrations of the first two fields reach 5-$\sigma$ limiting depths of $K_s=26.3$ (AB) and have seeing FWHM$\sim0\farcs4$.  Shorter MOSFIRE integrations of the latter two northern fields reach limiting depths $K_s=$\xxx\ with seeing FWHM=\xxx.  In all cases the $K_s$-band mosaics cover the primary cluster and parallel \textit{HST}/ACS+WFC3 fields, and the total area of the $K_s$-band coverage is \xxx\ arcmin$^2$.  The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$) and the IRAC 3.6\,\mum\ passband.  While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide significant constraints on both the redshifts and the stellar-population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).  Reduced, aligned mosaics of the four survey fields are provided accompanying this manuscript.  Time has been awarded for similar HAWKI coverage of the final two Frontier Fields Abell 370 and Abell S1063, with observations expected by the end of 2015.

\end{abstract}

\keywords{galaxies: evolution --- galaxies: high-redshift --- surveys: xxx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%   Introduction
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{s:introduction}

During the last two decades, near-infrared (NIR) imaging has gained a dominant role in studies of galaxy formation and evolution.

Detection of galaxies in the $K$-band ($\lambda_{\rm eff} \sim 2.2$~$\mu$m) provided the first opportunity to construct a comprehensive picture of the population of galaxies in the early universe, as it enabled the discovery of galaxies at $z > 2$ with suppressed emission in the optical (i.e., rest-frame UV) due to evolved stellar populations and/or significant amount of dust extinction (e.g., \textcolor{red}{refs}). In fact, these galaxies, which dominate the high-mass end of the high-$z$ galaxy population, were previously missed by rest-frame UV selection techniques (e.g., $U$-dropout galaxies; \textcolor{red}{refs}). Imaging in the $K$-band allows for the direct probing out to $z\approx5$ of rest-frame wavelengths longer than the Balmer break. Sampling the rest-frame optical wavelength regime is crucial for high-$z$ studies, as significantly less affected by dust obscuration and a better probe of the galaxy stellar mass compared to the rest-frame UV (more sensitive to unobscured star formation). While imaging at even longer wavelengths ($\lambda>3$~$\mu$m) are needed to probe the rest-frame optical of galaxies in the first billion years of cosmic history, deep sub-arsec resolution imaging at these wavelengths will not be available until the advent of the {\it James Webb Space Telescope} ({\it JWST}). Deep $K$-band data with good enough (i.e., FWHM$<1^{\prime \prime}$) image quality have been fundamental to fully exploit {\it Spitzer}-IRAC imaging data characterized by much poorer spatial resolution, allowing for the correction of the resulting blending of sources in IRAC images. Finally, for space-based studies, $K$ band crucially fills the gap between the reddest {\it Hubble Space Telescope} ({\it HST}) filter (i.e., F160W, with $\lambda \sim 1.6$~$\mu$m) and the IRAC 3.6~$\mu$m pass-band, greatly improving the constraints on both the redshifts and the stellar-population properties of galaxies.

It is not surprising that, given the aforementioned reasons, all successful extragalactic surveys performed in the last fifteen years to investigate galaxies populations in the early universe invested significant resources to obtain deep NIR imaging data with good image quality, especially in the $K$ band, usually adopted as the detection band. Indeed, our understanding of galaxy formation and evolution in the last couple of decades has impressively proceeded forward mostly driven by NIR progressively deeper, wider, and with better image quality, surveys, such as ...  FIRES, FIREWORKS, NMBS, UltraVISTA, etc... Subaru Super Deep Field (AO) \cite{minowa:05}... HUGS \citep{hugs}.\textcolor{red}{-- we should provide a complete list of NIR surveys here -- }

The latest effort to further our knowledge of galaxy formation and evolution is represented by the HST Frontier Fields program. \textcolor{red}{-- brief description of HFF, followed by the need for mega-deep K-band data --}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%   Observations
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observations \& Data Reduction}
\label{s:observations}

Table~\ref{tab:summary} provides a brief summary of the survey fields and observation characteristics.  Additional details on the observing strategy and image reduction procedures are provided in the subsections below.

\xxx\ currently a bit awkward since MOSFIRE reduction done quickly with Ivo's code and HAWK-I done in many additional steps (astrometry and photometry).

\subsection{Observation log}
\label{s:log}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%  Summary table
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{deluxetable*}{ccccrccc} 
\tabletypesize{\footnotesize}
\tablewidth{0pt} 
\tablecaption{Field summary \label{tab:summary}} 
\tablehead{ 
\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth$^\mathrm{a}$} & \colhead{FWHM}}
\startdata 
Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22 & 25.5 h & 25.9 & $0\farcs38$ \\ 
Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & 25.5 h & 25.9 & $0\farcs38$ \\ 
MACS-0717    & $07:17:34.0$ & $+37:44:49$ & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27, 2016 Jan 21--22  & 0/0 h\footnotemark[2] & 25.0/25.0 & $0\farcs43$/$0\farcs00$ \\ 
MACS-1149    & $11:49:36.3$ & $+22:23:58$ & \textit{Keck}/MOSFIRE  & 2015 Feb 24, 2016 Jan 21--22      & \xxx\footnotemark[2] & \xxx & \xxx \\
''           & ''           & ''          & \textit{VLT}/HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & \xxx & \xxx & \xxx
\enddata 
%\vspace{-0.8cm} 
\footnotetext[1]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
\footnotetext[2]{The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster/parallel \textit{HST} fields, which have the different characteristics as indicated, respectively.}
\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
\end{deluxetable*}
%\vspace{1cm}

$K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[HAWK-I;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).  HAWK-I $K_s$-badn bservations of the Abell S1063 and Abell 370 fields were obtained between 2015 Jul and 2016 Jan from a subsequent program (095.A-0533).  HAWK-I is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ HAWKI field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).

The HAWK-I observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally multiple fields were observed on the same night.  The execution of each OB lasted either 60 or 90 minutes, depending on how many exposures were obtained in the block.  The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744, 7$\times$8\,s~=~56\,s for MACS-0416, 7$\times$8\,s~=~56\,s for Abell S1063 and 7$\times$8\,s~=~56\,s for Abell 370\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since ESO Period 93.}).  The telescope was offset with random dithers between each exposure to facilitate sky subtraction;  observations taken before 2013 Nov 4 were dithered within a 20$^{\prime\prime}$ box, which was subsequently increased to 40$^{\prime\prime}$ to improve sky subtraction in the cluster cores crowded with bright galaxies and intra-cluster light.  The total on-sky integration times for the Abell 2744, MACS-0416, Abell S1063 and Abell 370 fields are 29.4, 25.5, \xxx, and \xxx\ hours, respectively.

Imaging observations of the MACS-0717 and MACS-1149 fields in the $K_s$ filter were obtained on 2015 Jan 26/27 and 2015 Feb 24 (program N097M), respectively, and both fields again on 2016 Jan 21/22 (program N135M), with the Multi-Object Spectrometer For Infra-Red Exploration \citep[MOSFIRE;][]{mosfire} on the 10\,m \mbox{\textit{Keck I}} telescope.  The MOSFIRE detector consists of $2048\times2048$ $0\farcs1798$ pixels; the resulting $6\farcm1\times6\farcm1$ field-of-view is unfortunately slightly too small to cover both \textit{HST} fields simultaneously and therefore requires two pointings to cover the entire deep \textit{HST} Frontier Fields imaging area (Fig.~\ref{fig:layout}).  The 6$\times$7\,s~=~42\,s MOSFIRE exposures were obtained in a $3\times3$ ``Box9'' dither pattern spaced roughly 40$^{\prime\prime}$ between offset positions.  The exposure times on the MOSFIRE survey fields are summarized in Table~\ref{tab:summary}.

Additional archival \mbox{HAWK-I} imaging of the MACS-1149 fields was obtained from the program 090.A-0458.  These data were obtained with one \mbox{HAWK-I} chip centered on the $z=9.7$ candidate from \cite{zheng:12}; they largely cover the HFF cluster field but the pointing was not optimized to also include the parallel field whose location was defined later (see Fig.~\ref{fig:layout}).  The MACS-1149 HAWK-I observations were obtained in Service Mode at the ESO/\textit{VLT} in \xxx\ separate OBs between 2013 Mar 21 and 2014 Jun 9.  The 7$\times$8\,s~=~56\,s \xxx\ exposures were taken at nine dithered positions with regular offsets separated by \xxx\ arcsec; the on-source exposure time of the final mosaic is \xxx\ hours.

\subsection{Image processing}
\label{s:processing}

The HAWK-I and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR \citep[zFOURGE;][, Labb\'e in prep]{spitler:12} infrared imaging instruments.  Treating each detector individually, the pipeline is easily modified for the different instrument configurations of the four HAWK-I and single MOSFIRE chips.  The primary task of the pipeline is removing the bright, time-variable sky background from the individual exposures, which is often some $10^4$ times brighter than the distant galaxies of interest in the field.  With such a bright background, we first determine an empirical ``sky flat'' that is a median of all of the \textit{science} exposures in a HAWK-I OB or MOSFIRE group, after rejecting the brightest 12 exposures at each pixel position to remove the contribution of bright objects.  We find these empirical flats to be preferable to external twilight or dome flats given the difficulty of obtaining a truly flat illumination pattern over such large detector fields-of-view.

After dividing by the flat, the background of each exposure is determined in a first pass from the simple median of the four exposures that came both immediately before and after it.  The background-subtracted exposures are combined into a mosaic from which objects are detected.  The final refined background of each exposure is determined in a second ``mask pass'' from a median again from the four exposures before and after it but now aggressively masking all pixels that contain flux from the detected objects.

As the archival HAWK-I images of the MACS-1149 field were obtained with longer individual exposures and with fewer exposures per sequence, the mask-pass technique described above did not produce satisfactory results.  In this case, for each raw exposure we divide by the sky flat and then subtract a third-order polynomial fit to the background.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Exposure Map figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\plotone{hawkiff_layout.pdf} % Figures/exposure_map.pdf
\caption{Layout of the Frontier Fields $K_s$-band mosaics.  The positions of the \textit{HST} cluster and parallel fields are shown in the blue (ACS optical) and red (WFC3 IR) polygons.  The light blue polygons in the MACS-0717 field show additional wide-field ACS imaging coverage from  programs GO-9722 and GO-10420 \citep[PI: Ebeling;][]{ma:11}.  The area covered by deep imaging in the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange.  The HAWK-I field of view is perfectly suited (Abell 2744, MACS-0416, Abell S1063 and Abell 370) for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE (MACS-0717 and MACS-1149).  Additional archival HAWK-I coverage of the MACS-1149 field is indicated with the gray square in the lower right panel.\label{fig:layout}}  
\end{figure*}

\subsection{Photometric calibration}
\label{s:calib}

For the HAWK-I observations, a single 90-minute Service Mode OB was obtained in each of the fields requiring photometric transparency conditions, and these OBs were followed immediately by an observation of a photometric standard star at a similar airmass with single exposures on each of the four chips.  The standard star exposures were processed with empirical sky flats derived from the science exposures as described above, and the observed fluxes of the standard stars yield absolute photometric zeropoints for each chip.  Correction factors were then computed to scale the OBs obtained at varying transparency levels to the calibrated photometric OB; the additional service mode OBs were obtained under generally satisfactory (i.e., ``clear'') weather conditions and the scale factors typically differ from unity by only a few percent.  The photometric calibration of the MOSFIRE observations was determined from standard-star observations taken the same night as the science exposures, again with the standard exposures reduced in the same way as the science exposures.

The 2-Micron All Sky Survey \citep[2MASS;][]{2mass} provides an additional check on the photometric calibration, though the comparison is limited due to relatively little brightness overlap between the faint end of reliable 2MASS photometry and the bright end where stars are in the linear regime of the detectors on the 8--10\,m telescopes.  A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Fig.~\ref{fig:2mass}.  The stellar photometry on the deep $K_s$-band mosaics is measured within $1\arcsec$ apertures corrected to infinity with the curves of growth described below.  The error-bars on the points in Fig.~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).  The 2MASS comparison suggests a systematic offset of $-0.09$~mag for the Abell S1063 field.  However, the exposure time for that field was similar to those of the additional deep HAWK-I fields and the derived depths of \textit{all} deep HAWK-I fields are nearly identical (Table~\ref{tab:summary}).  Given the overall agreement with the 2MASS photometry, we estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\leq0.05$~mag.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   2MASS comparison
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!t]
\epsscale{1.18}
\plotone{2mass_phot.pdf} % Figures/exposure_map.pdf
\caption{Comparison of bright-star photometry to the 2MASS public catalog.  The Abell 2744 and MACS 0416 fields at high galactic latitude have few stars for cross calibration, and the photometric calibration is determined from observations of standard stars on nights with photometric transparency conditions.  At relatively low galactic latitude ($b=+21$), the MACS 0717 field has many bright stars, and the photometric calibration in that field is defined from the 2MASS comparison (hence, $\Delta_\mathrm{2mass}\equiv0$, as indicated for that field).  We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}  
\end{figure}

\subsection{Astrometric alignment}
\label{s:astrometry}

In order to simplify measurements from the $K_s$-band images in consort with space-based \textit{HST} and \textit{Spitzer} mosaics, the astrometry of the ground-based mosaics must be refined.  Reference absolute astrometric catalogs were generated from \textit{HST} images when available (i.e., the chips overlapping the cluster and parallel \textit{HST} fields) and public \textit{Subaru} Suprime-Cam $r_c$-band images otherwise\footnote{http://archive.stsci.edu/pub/hlsp/clash/}.  Next, object catalogs are generated with the SEXtractor software \citep{bertin:96} for each individual background-subtracted exposure and transformations to the reference frame (shift, rotation and scale) are computed with with the \texttt{stsci.stimage.xyxymatch} and \texttt{scikit-image.transform}\footnote{\url{http://scikit-image.org}} Python software tools.  For the HAWK-I exposures, we fit a third-order polynomial to the geometric distortion model determined by \cite{libralato:14} and specify the polynomial terms as ``Simple Imaging Polynomial'' (SIP) coefficients \citep{fits:sip} in the individual exposure FITS files.  The SIP FITS header generally applicable for HAWK-I is provided in Appendix \xxx.  We note that the HAWK-I distortion is generally small, reaching $\sim$2 pixels ($0\farcs2$) at the image corners, but that the distortion corrections are necessary in particular given the excellent overall image quality of the observations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Image quality figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht]
\epsscale{1.18}
\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.  The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ and $\beta\sim2$, which have substantially more flux at radii $r>\mathrm{FWHM}$ than Gaussian profiles with the same half-widths (orange curves).  These extended profiles appear to be characteristic of deep $K_s$-band images; the stellar profiles from the HUGS GOODS-South D2 HAWK-I $K_s$ image \citep{hugs} (lower right panel), created by a different team with a different reduction pipeline, are nearly identical to those of the Frontier Fields mosaics presented here. \xxx\ I only show the parallel field region of MACS0717, which has significantly better image quality from the Jan. run.  You can see a few of the broader profiles in the figure; there are many more of these if you include the whole cluster area.  \label{fig:cog}}  
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%   Mosaic properties
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Drizzled $K_s$-band Mosaics}
\label{s:mosaic}

A crucial innovation of the image processing of the present analysis compared to previous works is that we combine all of the individual background-subtracted raw images (e.g., \xxx\ 4000 files for 1000 exposures times 4 detectors for the \mbox{Abell 2744} field) into the final output mosaic with the ``drizzle'' algorithm (\citealp{fruchter:02}, as implemented in the DrizzlePac/AstroDrizzle package; \citealp{drizzlepac}).  The benefits of the drizzle algorithm  preserving robust noise properties of the output mosaics will be discussed in detail below in \S\xxx.  Here we simply indicate that the drizzle implementation provides additional benefits in 1) trivial application of relative weights of the input images in creating the final mosaics, 2) easily applying the astrometric alignment and geometric distortion stored in the FITS headers, and 3) allowing the definition of an arbitrary output pixel grid without any additional image resampling.

Final mosaics of each field with $0\farcs1$ pixels are drizzled from the input exposures, $i$, with weights
\[
w^{-1} = F^2\cdot B \cdot \mathrm{FWHM}^\alpha / \sqrt{1-e^2}, 
\]
where $F$ is the factor to scale the exposure to the photometric system ($F\equiv1$ for the OB taken under photometric conditions, $F\gtrsim1$ otherwise), $B$ is the measured background, FWHM is the full width at half maximum of stars identified in the field, and $e$ is the ellipticity of the stellar point spread function (PSF).  The parameter $\alpha$ allows for optimizing the image quality of the final mosaic.  Increasing $\alpha$ puts larger weight on the exposures with the best image-quality while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.

Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies. The mosaics are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.  Inverse variance maps are shown in Appendix \xxx\.  The final image and inverse-variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).  The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity for faint galaxies.  In the sections below we describe the characteristics of the mosaics in more detail.

\subsection{Image quality}
\label{s:fwhm}

Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.  The image quality of the $K_s$-band mosaics is excellent, with stellar FWHM $0\farcs40$ and $0\farcs38$ in the Abell 2744 and MACS 0416 fields, respectively, thanks, in large part, to the execution of the HAWK-I observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.  Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.  Also shown in Fig.~\ref{fig:cog} are stellar curves of growth measured from the HAWK-I UDS and GOODS Survey \citep[HUGS;][]{hugs}, which are nearly identical in both FWHM and profile shape.  We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles. \xxx\ not clear what HAWK-I ETC assumes, the description just says ``FWHM of the seeing disk'' \xxx.  Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging than the redder \textit{Spitzer} IRAC bands.

\xxx\ add description of MOSFIRE clusters.  IQ not uniform across the MACS0717 field.

\subsection{Noise properties \& Depth}
\label{s:depth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Drizzle figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!t]
\epsscale{1.18}
\plotone{drizzle_cutouts.pdf} % Figures/exposure_map.pdf
\caption{Comparison of small regions of the final MACS 0416 mosaics with two output pixel sizes for images combined with the drizzle algorithm (top panels) and with the SWarp software (bottom panels).  With drizzle we can decrease the size of the input pixels before dropping them into the output grid, which significantly reduces the correlations between neighboring pixels.  Combining images with simple shift-and-add resampling such as with SWarp results in an effective smoothing of the pixel-to-pixel noise in the final combined image. 
\label{fig:drizzle_cutout}}  
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   correlated noise figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!t]
\epsscale{1.18}
\plotone{correlated_noise_figure.pdf} % Figures/exposure_map.pdf
\caption{The standard deviation of the pairwise pixel differences, sorted by pixel separation, provides a quantitative measure of the correlations between pixels.  The black curves determined from the drizzled images show very little variation with pixel separation, suggesting minimal correlations between adjacent pixels, as expected.  In contrast, the statistics of the SWarped images show a depression at small separations indicative of correlated pixels.  The magnitude and shape of the depression is reproduced exactly by the orange curves, which show the pairwise differences for the drizzled images convolved with small Gaussian kernels.  The effect is larger in the right panel, where the ratio of the input and output pixel sizes is larger by almost a factor of two.
\label{fig:corr_noise}}  
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Depth figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\epsscale{1.18}
\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
\caption{\textit{Top panel:} \xxx Noise in empty apertures as a function of aperture diameter, normalized to the smallest $D=0\farcs2$ aperture.  The dashed and dotted lines show the expectations for uncorrelated and perfectly-correlated noise within the aperture.  \textit{Bottom panel:} The product of the inverse curves-of-growth (Fig.\ref{fig:cog} and the empty aperture noise gives the point-source sensitivity as a function of aperture size.  The S/N is maximized with an aperture just larger than the FWHM.  For reference, the colored dashed lines in the bottom panel indicate the depth computed assuming uncorrelated noise within the apertures ($\sigma \propto D$) and the per-pixel variance reported in the weight maps.)
\label{fig:depth}}  
\end{figure}
%%%%%%

Along with practical benefits mentioned above, the most dramatic benefit of using the drizzle algorithm comes from the fact that the raw detector pixels are only resampled once in the process of generating the output mosaic.  Furthermore, with a large number of dithered raw images we can use drizzle to shrink the raw input pixels by a factor of ten before dropping them into the output grid,\footnote{The pixels are shrunk by adopting pixfrac=0.1 as defined by \cite{fruchter:02}.} and the result is dramatically reduced correlations between neighboring pixels, particularly for output pixel sizes that are somewhat smaller than the native HAWK-I or MOSFIRE detector pixels.

A comparison of final combined images created with output $0\farcs1$ and $0\farcs06$ pixel grids using the drizzle implementation to versions created with the SWarp software \citep{swarp} is shown in Fig.~\ref{fig:drizzle_cutout}.  In the later case, the raw pixels are resampled twice before making the final mosaic, once combining the raw exposures in each OB and a second time combining the separate OBs into the final stack with an arbitrary pixel grid.  One could avoid the second resampling and combine all of the raw exposures at once, though SWarp is still limited to dropping the original pixels into the output mosaic preserving the input native pixel grid and therefore a single input pixel maps onto multiple (perhaps many) output pixels.

The fact that the drizzle-combined images in Fig.~\ref{fig:drizzle_cutout} (panels \textit{a} and \textit{b}) appear to the eye to be noisier than their SWarped counterparts (panels \textit{c} and \textit{d}) is simply the result of a smoothing of the pixel noise in the latter case rather than reflecting true underlying differences in their pixel-to-pixel variance.  \cite{casertano:00} discuss how the noise properties of correlated output pixels are affected by the relative sizes of the input and output pixel grids and the adopted drizzle parameters.  With the drizzle algorithm and a small pixfrac, a given input pixel will map to only a single output pixel and therefore correlations between the output pixels should be minimal\footnote{Inter-pixel capacitance and other forms of crosstalk intrinsic to the IR detectors effectively place a lower limit on the correlations between adjacent pixels \citep{hilbert:ipc}.}.

We explore these correlations between the output pixels in Fig.~\ref{fig:corr_noise}.  For the black and red curves in each panel, we measure value differences between random pairs of pixels sorted by the separation between the pixels for the drizzled and SWarped images, respectively.  For perfectly uncorrelated pixels, these curves will be flat and will reflect the intrinsic noise of the image pixels.  In the presence of correlations between adjacent pixels, however, the curves will show a depression at small separations.  As expected, the SWarp-combined images show just such a depression for pairs of pixels 1--2 pixels apart.  The yellow curves in Fig.~\ref{fig:corr_noise} show the same pairwise differences for the drizzled images now convolved by small Gaussian kernels as indicated, which agree very well with the curves for the images with correlated pixels.

In terms of understanding the noise properties of the images, a key point here is not only that depression at small separations indicates the presence of correlations between adjacent pixels but also the fact that correlations cause the apparent overall r.m.s. to be decreased at all separations.  That the pixel variances can't be trusted in the presence of such correlations is the primary reason why significant effort has been devoted to placing random ``empty apertures'' across images in order to characterize their noise properties \citep[e.g., ]{quadri:musyc, whitaker:nmbs, skelton:14}.  Now by eliminating the pixel correlations with the drizzle combination (with a small pixfrac) we have shown that the pixel statistics are robust and it is now trivial to compute the expected variance within an arbitrary photometric aperture, for example:
\begin{equation}
\sigma_D^2 = \sigma_\mathrm{pix}^2\cdot \pi/4 \cdot D^2,
\label{eq:aper_var}
\end{equation}
for circular apertures with diameter, $D$, in pixels, and where $\sigma_\mathrm{pix}^2$ is the per-pixel variance determined from an analysis such as that in Fig.~\ref{fig:corr_noise}.

The product of Eq.~\ref{eq:aper_var} and the inverse of the curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.  For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs6$, just larger than the FWHM of point-source profiles.  The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are \xxx\ 25.8, 25.9, and 25.0 in the Abell 2744, MACS-0416, and MACS-0717 fields, respectively.  The HAWK-I images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.

We conclude here with a brief comment on the ``HUGS'' ultra-deep HAWK-I images of the CANDELS survey fields.  \cite{hugs} report a final depth of 26.5 AB for point sources within $D=0\farcs4$ apertures for their deepest field, ``GOODS-D1'', which has an exposure time of 31.5 hours similar to the deep HAWK-I fields summarized in Table~\ref{tab:summary}.  \cite{hugs} assume ideal noise properties that likely suffer some residual correlations between adjacent pixels such as that indicated by the red curves in Fig.~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.  This effect ($\sim$0.4 mag) along with the relative exposure times ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields HAWK-I images described here.  The comparison must be made because any additional half magnitude increase depth is exponentially more expensive to obtain!

\xxx\ say something about distant extended sources, perhaps scaling from Arjen's size distributions?

\subsection{Number counts}
\label{s:number}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Cutout figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\epsscale{1.18}
\plotone{image_cutouts.pdf} % Figures/exposure_map.pdf
\caption{Image cutouts of the MACS 0416 parallel field.  The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $10\arcsec$ \xxx\ [check] on a side.
\label{fig:cutout}}  
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Number counts figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\epsscale{1.18}
\plotone{number_counts_all.pdf} % Figures/exposure_map.pdf
\caption{$K_s$-band number counts in the Frontier Fields. \label{fig:counts}}  
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%   Constraints
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Photometric constraints from the $K_s$ band}
\label{s:constraints}

\subsection{Photometric redshifts}
\label{s:photoz}

\subsection{SED modeling}
\label{s:sed}

Balmer break, UV slope

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%   Discussion and Summary
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and Conclusions}
\label{s:discussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   End things
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-0.3cm} %% emulateapj
\acknowledgements

\noindent We gratefully acknowledge funding support from the STScI Director's Discretionary Research Fund.  \xxx\ acknowledge NASA/Keck time + Mauna Kea \xxx. This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy}, and open-source Python modules Numpy, Scipy, and Matplotlib.

% \noindent We gratefully acknowledge funding support from the Marie Curie Actions of the European Commission (FP7-COFUND) and STScI
% grant GO-12177.

{\it Facilities:} \facility{VLT:Yepun (HAWKI)}, \facility{Keck:I (MOSFIRE)} % \facility{Hubble Space Telescope (WFC3)}

\bibliographystyle{apj_url}
\bibliography{ms}

\end{document}